Canonical rings of $\bq$-divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
S's changed to T's
fractional/rational divisors -> Q-divisor

\title{Canonical rings of rational divisors on projective space and Hirzebruch surfaces}

@@

Another closely related topic is the Hassett-Keel program 
initiated in \todo{was it actually initiated here? check vzb.} \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0}H^0 \left( \overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} + \alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces. Fractional divisors on surfaces also naturally appear when considering the canonical ring of surfaces of Kodaira dimension 1. A surface has Kodaira dimension 1 if and only if it is an elliptic surface, and their canonical ring are most naturally described in the setting of fractional divisors. \todo{Add references}

@@

\noindent
where
\[
	\tau = \sigma
	+ \max \left( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{
	\substack{i \in	S_+ \\ j\in S_-}}(\ell_{i,j} (\bida_i \bidb_j
	- \bida_j \bidb_i))	\right).
\]
\end{thm}

@@

\ssec{Notation}
\begin{convention}
Notate a fractional (Weil) divisor
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes \bq.
\end{align*}
where $\alpha_i \in \bq$ and $D_i \in \di X$ is an irreducible divisor.
\end{convention}

@@

\begin{lem}
$\ker(\psi)$ is generated in less than
\[
	\tau = \sigma
	+ \max \left( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{\substack{
	i \in	S_+ \\ j\in S_-}} (\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))
	\right).
\]
\end{lem}

@@

\noindent
where $E$ is $\mathbb{R}_{\ge 0}$
combination of elements in 
\[
	\{e_i \mbox{ if } i \in S\} \cup \{e_{i, j}: j \in S_- \mbox{ if }
	i	\in S_+ \} \cup \{e_{j, i}: j \in S_+ \mbox{ if } i \in S_- \}.
\]

@@

\noindent
which divides $(z_i - \beta_i) \prod_{i = 1}^n y_i^{c_i}$.
We can bound $\deg(F)$ by
\[
	\deg(F) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in S_= \\
	\max_{j \in S_-} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in S_+ \\
	\max_{j \in S_+} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in S_- \end{cases}.
\]

\noindent
Therefore, $\ker(\psi)$ is generated in degree at most
\[
	\tau = \sigma
	+ \max \left(\max_{i\in S_=} \bigl(\ell_i \gcd(a_i, b_i) \bigr),
	\; \max_{\substack{i \in S_+ \\ j \in S_-}} \bigl(\ell_{i, j}
	(\bida_i \bidb_j - \bida_j
	\bidb_i) \bigr) \right).
\]
\end{proof}

@@

\noindent
where
\[
	\tau = \sigma
	+ \max \left( \max_{i\in S_=} \bigl(\ell_i \gcd(a_i, b_i) \bigr),
	\max_{\substack{i \in S_+ \\ j \in S_-}} \bigl(\ell_{i, j}
	(\bida_i \bidb_j - \bida_j \bidb_i) \bigr) \right).
\]
\end{thm}

>>>>

\title{Canonical rings of $\bq$-divisors on projective space and Hirzebruch surfaces}

@@

Another closely related topic is the Hassett-Keel program 
initiated in \todo{was it actually initiated here? check vzb.} \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0}H^0 \left( \overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} + \alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces. $\bq$-divisors on surfaces also naturally appear when considering the canonical ring of surfaces of Kodaira dimension 1. A surface has Kodaira dimension 1 if and only if it is an elliptic surface, and their canonical ring are most naturally described in the setting of $\bq$-divisors. \todo{Add references}

@@

\noindent
where
\[
	\tau = \sigma
	+ \max \left( \max_{i\in \Te}(\ell_i \gcd(a_i, b_i)), \max_{
	\substack{i \in	\Tp \\ j\in \Tm}}(\ell_{i,j} (\bida_i \bidb_j
	- \bida_j \bidb_i))	\right).
\]
\end{thm}

@@

\ssec{Notation}
\begin{convention}
Notate a (Weil) $\bq$-divisor
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes \bq.
\end{align*}

\noindent
where $\alpha_i \in \bq$ and $D_i \in \di X$ is an irreducible divisor.
\end{convention}

@@

\begin{lem}
$\ker(\psi)$ is generated in less than
\[
	\tau = \sigma
	+ \max \left( \max_{i\in \Te}(\ell_i \gcd(a_i, b_i)), \max_{\substack{
	i \in	\Tp \\ j\in \Tm}} (\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))
	\right).
\]
\end{lem}

@@

\noindent
where $E$ is $\mathbb{R}_{\ge 0}$
combination of elements in 
\[
	\{e_i \mbox{ if } i \in S\} \cup \{e_{i, j}: j \in \Tm \mbox{ if }
	i	\in \Tp \} \cup \{e_{j, i}: j \in \Tp \mbox{ if } i \in \Tm \}.
\]

@@

\noindent
which divides $(z_i - \beta_i) \prod_{i = 1}^n y_i^{c_i}$.
We can bound $\deg(F)$ by
\[
	\deg(F) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]

\noindent
Therefore, $\ker(\psi)$ is generated in degree at most
\[
	\tau = \sigma
	+ \max \left(\max_{i\in \Te} \bigl(\ell_i \gcd(a_i, b_i) \bigr),
	\; \max_{\substack{i \in \Tp \\ j \in \Tm}} \bigl(\ell_{i, j}
	(\bida_i \bidb_j - \bida_j
	\bidb_i) \bigr) \right).
\]
\end{proof}

@@

\noindent
where
\[
	\tau = \sigma
	+ \max \left( \max_{i\in \Te} \bigl(\ell_i \gcd(a_i, b_i) \bigr),
	\max_{\substack{i \in \Tp \\ j \in \Tm}} \bigl(\ell_{i, j}
	(\bida_i \bidb_j - \bida_j \bidb_i) \bigr) \right).
\]
\end{thm}
